---
title: "HW2_SERCAN_BÖÇKÜN_2019402123"
author: "Sercan Böçkün"
date: '2022-05-07'
output: html_document
---



Installing required libraries and preparing the data

```{r setup, echo=TRUE}
require(data.table)

require(lubridate)
require(forecast)
require(ggplot2)
require(GGally)


############# Reading and processing datas ##########
data_path="C:/Users/serca/OneDrive/Masaüstü/Spring 22'/IE 360/HW_2/IE360_Spring22_HW2_data.csv"
data=read.csv(data_path)

data$Unleaded.Gasoline.Sale..UGS. = gsub( " ", "", data$Unleaded.Gasoline.Sale..UGS. )
data$Unleaded.Gasoline.Sale..UGS. = as.numeric(data$Unleaded.Gasoline.Sale..UGS.)

data$X..LPG.Vehicles..NLPG. = gsub( " ", "", data$X..LPG.Vehicles..NLPG. )
data$X..LPG.Vehicles..NLPG. = as.numeric(data$X..LPG.Vehicles..NLPG)

data$X..Unleaded.Gasoline.Vehicles..NUGV. = gsub( " ", "", data$X..Unleaded.Gasoline.Vehicles..NUGV. )
data$X..Unleaded.Gasoline.Vehicles..NUGV. = as.numeric(data$X..Unleaded.Gasoline.Vehicles..NUGV.)

data$GNP.Agriculture = gsub( " ", "", data$GNP.Agriculture )
data$GNP.Agriculture = as.numeric(data$GNP.Agriculture)

data$GNP.Commerce = gsub( " ", "", data$GNP.Commerce )
data$GNP.Commerce = as.numeric(data$GNP.Commerce)

data$GNP.Total = gsub( " ", "", data$GNP.Total )
data$GNP.Total = as.numeric(data$GNP.Total)


data$Quarter = as.character(data$Quarter)
#data$Quarter <- as.Date(data$Quarter)  

data = data.table(data)

head(data,25)
str(data)
```

## 0. Introduction

In this document statistical analysis and regression models will be discussed and applied to the dataset which was presented in the
previous section. The last 4 rows of the "Unleaded.Gasoline.Sale..UGS." column is missing and represents the unleaded gasoline sales in a given quarter.


## 1. Plotting and Commenting on the UGS 


```{r 1, echo=TRUE}
data$Unleaded.Gasoline.Sale..UGS. = ts(data$Unleaded.Gasoline.Sale..UGS., start = 2000, deltat= 1/4)

ggplot(data, aes(Quarter,Unleaded.Gasoline.Sale..UGS. , group=1)) + geom_line() +geom_point()
```

Looking at the shape of the plot it can be observed that the number of sales have a decreasing trend with seasonal peaks and lows. Visiting back to the definition of stationary we know that stationary time series data is not predictable since it has no any detectable trend and seasonality. Our UGS data is clearly shrinking with a slight decrase in it's variance. We conclude that the data is not stationary thus prediction models can be prepared at tested.

## 2. Looking at the Autocorrelation of UGS

```{r 2, echo=TRUE}
acf(data$Unleaded.Gasoline.Sale..UGS.[0:28], 12) # The last 4 values are missing. We use the available data to measure autocorrelation func.
```

Above plotting shows the autocorrelation of the UGS sales. One can easily detect that ACF is having peaks with lags that are multiplier of 4. This is a indication of seasonality with respect to quarters. 


## 3. Adding Seasonality and Trend data

```{r 3, echo=TRUE}
data[,trend:=1:.N] # Adding the trend factor as a column to the data

data[,seasonality:=((1:.N)%%4)] # remainder from 4. 4th quarters are represented with "0".
data[seasonality == 0, seasonality := 4]
data$seasonality = as.factor(data$seasonality) # Taking seasonality as factor

str(data)

```



## 4. Building Regression Models

Before starting to building models let's take a look at the correlation matrix to have a better understanding of the data.
```{r 4, echo=TRUE}

ggpairs(data[,-1,with=FALSE]) # We will build our model with the variables which are in high correlation with our output variable 
```


Base Model:
```{r 40, echo=TRUE}
model_0= lm(Unleaded.Gasoline.Sale..UGS.~trend ,data)
summary(model_0)
checkresiduals(model_0$residuals) # There's high autocorrelation with lag 2 and 4's

```

Adding seasonality and trend to the base model.
```{r 41, echo=TRUE}
model_1= lm(Unleaded.Gasoline.Sale..UGS.~trend + seasonality,data) # One of the seasonality is omitted into the intercept
summary(model_1)
checkresiduals(model_1$residuals)

```

Let's start to add the variables to the model respectively with their correlation with our output variables to increase the performance (Adj R^2) of the model and drop the insignificant variables.


```{r 42, echo=TRUE}
model_2= lm(Unleaded.Gasoline.Sale..UGS.~trend + seasonality 
            + Price.of.Diesel.Gasoline..PG.+
              X..Unleaded.Gasoline.Vehicles..NUGV. + X..of.Diesel.Gasoline.Vehicles..NDGV.
             + GNP.Commerce , data)
#+X..Unleaded.Gasoline.Vehicles..NUGV. + X..LPG.Vehicles..NLPGGNP.Total,data)
summary(model_2)
checkresiduals(model_2$residuals)
checkresiduals(model_2) # 
```

Our adjusted R^2 increased but there's still a high autocorrelation in our residuals with a lag of one. Let's add y(t-1) and y(t-4) to the data

```{r 421, echo=TRUE}
setDT(data)

data[,.(one_lagged_sales = shift(data$Unleaded.Gasoline.Sale..UGS., n=1L, fill=NA))]
data$one_lagged_sales = data[,.(two_lagged_sales = shift(data$Unleaded.Gasoline.Sale..UGS., n=1L, fill=NA))]

data[,.(four_lagged_sales = shift(data$Unleaded.Gasoline.Sale..UGS., n=4L, fill=NA))]
data$four_lagged_sales = data[,.(four_lagged_sales = shift(data$Unleaded.Gasoline.Sale..UGS., n=4L, fill=NA))]
```

We added the 4 and 1 time interval lagged y(t)'s to their corresponding rows. The previous model had a high autocorrelation with a lag of 1. Let's add one_lagged_sales to the model

```{r 43, echo=TRUE}
model_3= lm(Unleaded.Gasoline.Sale..UGS.~trend + seasonality 
            + Price.of.Diesel.Gasoline..PG.+
              X..Unleaded.Gasoline.Vehicles..NUGV. + X..of.Diesel.Gasoline.Vehicles..NDGV.
            + GNP.Commerce + one_lagged_sales  , data)
summary(model_3)
checkresiduals(model_3$residuals)
```


The high autocorrelation with lag 1 is ommitted with the addition of 1 lagged output data ACF values are currently below the upper and lower limits. However significance of other variables reduced. Let's drop insignificant variables from the model except the seasonality. Even though seasonality variable is insignificant, when it's reduced there occurs spikes in autocorrelation of residuals with period of 4 despite the little increase in the Adj R^2. For that reason we will keep it in the model for the sake of normality of our residuals. 

```{r 44, echo=TRUE}
model_4= lm(Unleaded.Gasoline.Sale..UGS.~trend + seasonality 
            + Price.of.Diesel.Gasoline..PG.+
              X..Unleaded.Gasoline.Vehicles..NUGV. + X..of.Diesel.Gasoline.Vehicles..NDGV.
          + one_lagged_sales  , data)
summary(model_4)
checkresiduals(model_4$residuals)
```

We have dropped the  "GNP.Commerce" variable since it was found to be insignificant from the t test. Looking at the residuals and it's autocorrelation, we still see some negative autocorrelation with lag of 4. We might consider to add four_lagged_sales to the model. 

```{r 45, echo=TRUE}
model_5= lm(Unleaded.Gasoline.Sale..UGS.~trend + seasonality 
            + Price.of.Diesel.Gasoline..PG.+
              X..Unleaded.Gasoline.Vehicles..NUGV. + X..of.Diesel.Gasoline.Vehicles..NDGV.
            + one_lagged_sales + four_lagged_sales , data)
summary(model_5)
checkresiduals(model_5$residuals)
```


We have added the four_lagged_sales to the model in order to reduce the autocorrelation with the lag of 4 but contrary to our expectation this variable is found to be insignificant from the t test and our model's adjusted R^2 is reduced heavily. In conclusion we will select the model_4 as our main model.


## 5. Final Model
```{r 5, echo=TRUE}
summary(model_4)
checkresiduals(model_4$residuals)
```

Our final model's residual's autocorrelation values are within the decided upper and lower limits except the slight breach of the lower limit at the lag 4. Model's Adj R^2 is 0.9651 and residuals are distribution can be estimated to be normal. Most of our variables are mostly significant and our least significant variable has a p value of 0.088 which can be considered somewhat significant  


## 6. Making Predictions
The time to make predictions for the quarters in 2007 has come. Before diving deep into the predictions one thing should be clarified.
Predicting the sales in the rows of 2007_Q2, 2007_Q3, 2007_Q4 is challenging since these rows do not have the data of "previous year's sales", these data points are blank and model cannot do predictions for these data points. We will put in model's sales prediction for the "2007_Q1" as the "previous year's sales" data point to the "2007_Q2", "2007_Q2" to the "2007_Q3" and "2007_Q3" to the "2007_Q4"
```{r 6, echo=TRUE}
tmp=copy(data)

tmp[,actual:=Unleaded.Gasoline.Sale..UGS.]
tmp[,predictions:=predict(model_4,tmp)]
print(tmp)


tmp[ 30, "one_lagged_sales"] = 66210.0
tmp[,predictions2:=predict(model_4,tmp)]


tmp[ 31, "one_lagged_sales"] = 1110417.4
tmp[,predictions2:=predict(model_4,tmp)]


tmp[ 32, "one_lagged_sales"] = 848645.6
tmp[,predictions2:=predict(model_4,tmp)]

```



## 7. Predictions
```{r 7, echo=TRUE}
print(tmp$predictions2)

ggplot(tmp ,aes(x=Quarter)) +
  geom_line(aes(y=actual,color='real', group = 1)) + 
  geom_line(aes(y=predictions2, color = 'predictions', group = 1) ) 

```





